---
title: "ETC5513_Assignment1"
author: "Sinping CHAN"
editor: visual
format:
  html:
    theme: cosmo       
    toc: true
    embed-resources: true 
  pdf:
    documentclass: scrartcl
    geometry: margin=0.8in
    toc: true
    number-sections: true
    colorlinks: true
execute:
  echo: false         
  warning: false
  message: false
---

## [Introduction]{.underline}

In an increasingly competitive e-commerce landscape, quantifying the return on personalized marketing investments is essential for driving sustainable revenue growth. This research examines whether higher expenditure on advertisements, emails, and coupons during the previous quarter’s personalized campaign led to increased individual purchase amounts. By analyzing these relationships at the customer level, the study evaluates the effectiveness of variable marketing spend across different channels.

## [Research Question]{.underline}

This analysis explores whether increased marketing investment per customer effectively drives higher purchase revenue. By evaluating the relationship between spending and performance across various channels, this report identifies whether current strategies yield linear growth or if specific segments have reached a point of diminishing returns.

## [Dataset Introduction]{.underline}

```{r}
#| label: data-cleansing
#| warning: false
#| message: false
#| echo: false
#| results: "hide"

# ============================================================
#  Data quality issues addressed:
#    1. Mixed date formats  → standardised to ISO 8601 (YYYY-MM-DD)
#    2. Missing signup_date  (9 rows)  → kept as NA
#    3. Missing purchase_amount (18 rows) → kept as NA; flag added
#    4. notes column: NaN → NA; meaningful values retained
#    5. segment → ordered factor (Consumer < SMB < Enterprise)
#    6. Derived features: days_since_signup, roi, contacted_flag,
#                         returned_flag, has_purchase
# ============================================================

# ── 0. PACKAGES ─────────────────────────────────────────────
library(dplyr)
library(lubridate)
library(stringr)
library(readr)

# ── 1. LOAD RAW DATA ────────────────────────────────────────

raw <- read_csv(
  "data/35851937_Sin Chan.csv",   # adjust path if needed
  col_types = cols(
    customer_id      = col_character(),
    signup_date      = col_character(),   # read as text; parsed below
    marketing_spend  = col_double(),
    visits           = col_integer(),
    purchase_amount  = col_double(),
    segment          = col_character(),
    notes            = col_character()
  ),
  na = c("", "NA", "N/A", "na", "n/a")   # normalise common NA strings on read
)

cat("Raw data loaded:", nrow(raw), "rows x", ncol(raw), "cols\n")


# ── 2. DATE PARSING HELPER ──────────────────────────────────
# The column contains (at least) four formats:
#   YYYY-MM-DD     e.g. "2025-06-24"
#   DD/MM/YYYY     e.g. "22/05/2025", "27/10/2024"
#   MM/DD/YYYY     e.g. "05/10/2024"  ← ambiguous; resolved via context*
#   Mon DD, YYYY   e.g. "Feb 05, 2025", "Oct 10, 2024"
#
# * DD/MM vs MM/DD is ambiguous when day <= 12. Inspection of the data
#   confirms the supplier locale uses DD/MM/YYYY throughout; that format
#   is tried before MM/DD/YYYY so it takes precedence.

parse_mixed_date <- function(x) {
  formats <- c(
    "%Y-%m-%d",    # ISO  – try first (most common)
    "%d/%m/%Y",    # DD/MM/YYYY
    "%m/%d/%Y",    # MM/DD/YYYY (fallback)
    "%b %d, %Y",   # Mon DD, YYYY  (abbreviated month)
    "%B %d, %Y"    # Month DD, YYYY (full month name)
  )

  result <- as.Date(rep(NA_character_, length(x)))

  for (fmt in formats) {
    needs_filling <- is.na(result) & !is.na(x)
    if (!any(needs_filling)) break
    parsed <- suppressWarnings(as.Date(x, format = fmt))
    result[needs_filling & !is.na(parsed)] <- parsed[needs_filling & !is.na(parsed)]
  }
  result
}


# ── 3. CLEAN ────────────────────────────────────────────────

reference_date <- as.Date("2026-03-31")   # today; used for derived fields

cleaned <- raw %>%

  # 3a. Parse signup_date
  mutate(
    signup_date = parse_mixed_date(signup_date)
  ) %>%

  # 3b. notes: already NA-normalised on read; standardise case
  mutate(
    notes = str_to_lower(str_trim(notes))   # "Contacted" → "contacted"
  ) %>%

  # 3c. segment → ordered factor
  mutate(
    segment = factor(
      segment,
      levels  = c("Consumer", "SMB", "Enterprise"),
      ordered = TRUE
    )
  ) %>%

  # 3d. Derived features
  mutate(
    # Days between signup and reference date (NA if signup unknown)
    days_since_signup = as.integer(reference_date - signup_date),

    # Return on marketing investment; NA if either value is missing
    roi = ifelse(
      !is.na(purchase_amount) & marketing_spend > 0,
      round((purchase_amount - marketing_spend) / marketing_spend, 4),
      NA_real_
    ),

    # Boolean flags derived from notes
    contacted_flag  = (!is.na(notes) & notes == "contacted"),
    returned_flag   = (!is.na(notes) & notes == "returned"),
    no_response_flag = (!is.na(notes) & notes == "no response"),

    # Flag rows with missing purchase_amount for downstream imputation decisions
    purchase_missing = is.na(purchase_amount)
  )


# ── 4. VALIDATION CHECKS ────────────────────────────────────

cat("\n", rep("=", 60), "\n")
cat(" DATA QUALITY & PERFORMANCE METRICS SUMMARY \n")
cat(rep("=", 60), "\n\n")

# 1. 數據基礎規模
cat(sprintf("%-30s : %d\n", "Total Customer Records", nrow(cleaned)))
cat(sprintf("%-30s : %d\n", "Total Analysis Variables", ncol(cleaned)))
cat("\n")

# 2. Key Performance Metrics
# Mean, Median and SD to show the distribution of the data
metrics_summary <- cleaned %>%
  summarise(
    # Marketing Expenditure
    mean_spend    = mean(marketing_spend, na.rm = TRUE),
    median_spend  = median(marketing_spend, na.rm = TRUE),
    
    # Total Purchase Amounts
    mean_purchase = mean(purchase_amount, na.rm = TRUE),
    sd_purchase   = sd(purchase_amount, na.rm = TRUE),
    
    # Platform Visit Frequency
    mean_visits   = mean(visits, na.rm = TRUE),
    max_visits    = max(visits, na.rm = TRUE),
    
    # ROI
    mean_roi      = mean(roi, na.rm = TRUE)
  )

cat("Performance Metrics (Averages & Distribution):\n")
cat(sprintf("  %-28s : $%.2f (Median: $%.2f)\n", "Avg Marketing Expenditure", metrics_summary$mean_spend, metrics_summary$median_spend))
cat(sprintf("  %-28s : $%.2f (Std Dev: %.2f)\n", "Avg Purchase Amount", metrics_summary$mean_purchase, metrics_summary$sd_purchase))
cat(sprintf("  %-28s : %.1f visits (Max: %d)\n", "Avg Visit Frequency", metrics_summary$mean_visits, metrics_summary$max_visits))
cat(sprintf("  %-28s : %.1f%%\n", "Average ROI", metrics_summary$mean_roi * 100))

# 3. 數據完整性檢查 (Missing Data)
cat("\nData Integrity Check (Missing Values):\n")
missing_counts <- colSums(is.na(cleaned))
# List out all missing value
significant_missing <- missing_counts[missing_counts > 0]
if(length(significant_missing) > 0) {
  for(i in seq_along(significant_missing)) {
    cat(sprintf("  %-28s : %d rows missing\n", names(significant_missing)[i], significant_missing[i]))
  }
} else {
  cat("  All key columns are 100% complete.\n")
}

# 4.Segment
cat("\nMarket Segment Distribution:\n")
seg_table <- table(cleaned$segment)
for(n in names(seg_table)) {
  cat(sprintf("  %-28s : %d customers\n", n, seg_table[n]))
}

cat("\n", rep("=", 60), "\n")

# ── 5. EXPORT ───────────────────────────────────────────────

write_csv(cleaned, "data/35851937_Sin_Chan_cleaned.csv", na = "")

```

### Data Overview

This dataset records customer engagement and marketing performance of `r nrow(cleaned)` customers across 3 market segments (Consumer, SMB, and Enterprise) from September 2024 to February 2026. The data measures key performance metrics including marketing expenditure, purchase amounts, visit frequency, and return on investment (ROI).

By analyzing these variables, the report aims to evaluate the effectiveness of personalized marketing spend and identify optimal investment windows for future campaign cycles.

### Data Cleaning

To ensure the dataset was ready for reproducible analysis, a cleaning procedure was performed. First, the `signup_date` column was standardized into a consistent ISO 8601 format to resolve various date string discrepancies. Second, inconsistent missing value indicators (such as 'N/A' or 'na') were unified into standard R `NA` values across all columns, particularly for `purchase_amount`. This standardization was critical for the derivation of the Return on Investment (ROI) metric; by maintaining these 18 missing purchase records as `NA`, the subsequent analysis avoids skewed results that would occur if missing values were incorrectly treated as zero.

Textual entries in the `notes` column were trimmed of extra whitespace and converted to lowercase to maintain string consistency. Furthermore, the `segment` variable was reclassified as an ordered factor to correctly represent the hierarchy of business tiers (Consumer \< SMB \< Enterprise). This ensures that both the ROI summary statistics and the LOESS smoothing curves reflect the intended business logic while accounting for the inherent gaps in the transactional data.

## [Dataset Description]{.underline}

```{r}
#| label: data-details
#| warning: false
#| message: false
#| echo: false
#| results: "hide"

library(skimr)

# A high-level description of all variables
# This will show you means, standard deviations, and distributions
skim(cleaned)
```

The processed dataset consists of `r nrow(cleaned)` observations and `r ncol(cleaned)` variables, ranging from numerical financial figures to categorical classifications and logical flags used for tracking customer behavior. Notably, some variables have missing values, particularly in signup_date and notes, which may affect data completeness and the reliability of derived measures.

### Variable Classifications

The dataset captures customer behavior through six distinct data types:

-   **Numeric**: `marketing_spend`, `purchase_amount`, and `roi`.

-   **Integer**: `visits` and `days_since_signup`.

-   **Ordered Factor**: `segment` (Consumer \< SMB \< Enterprise).

-   **Date**: `signup_date`.

-   **Character**: `customer_id` and `notes`.

-   **Logical**: `contacted_flag`, `returned_flag`, `no_response_flag`, and `purchase_missing`

### Segment Composition and Data Integrity

The sample is predominantly composed of the Consumer segment (n=94), followed by SMB (n=37) and Enterprise (n=19) tiers. While the dataset is comprehensive, 12% of the observations (n=18) lack purchase and ROI data. These records are explicitly flagged as missing to ensure that subsequent non-linear modeling remains robust and transparent.

### Preliminary Descriptive Analysis

```{r}
#| label: tbl-segment-summary
#| echo: false
#| message: false

library(knitr)

segment_summary <- cleaned |>
  group_by(segment) |>
  summarise(
    `Avg Spend` = mean(marketing_spend, na.rm = TRUE),
    `Avg Purchase` = mean(purchase_amount, na.rm = TRUE),
    `Total Visits` = sum(visits, na.rm = TRUE),
    `Missing Rate (%)` = mean(purchase_missing) * 100
  ) |>
  mutate(across(where(is.numeric), ~round(., 2)))

kable(segment_summary, 
      caption = "Summary Statistics of Campaign Performance by Segment")


```

```{r}
#| label: fig-spend-vs-purchase
#| fig-cap: "Marketing Spend vs Purchase Amount by Segment"
#| echo: false
#| warning: false

library(ggplot2)

ggplot(cleaned, aes(x = marketing_spend, y = purchase_amount, color = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Marketing Spend ($)",
    y = "Purchase Amount ($)",
    color = "Segment"
  ) +
  theme_minimal()

```

\
As detailed in @tbl-segment-summary, the average marketing spend and purchase behavior vary across the three segments, with the Enterprise segment showing the highest average marketing spend at \$174.73, while the Consumer segment contributed the vast majority of platform traffic with 511 total visits.\

Furthermore, @fig-spend-vs-purchase illustrates the linear relationship between spending and revenue, a deeper look at the ROI reveals a more complex financial reality. Although the mean ROI is 12.7%, the typical (median) ROI is slightly negative at -7%. This highly skewed distribution suggests that a small number of high-performing accounts offset a larger group of underperforming customers who have yet to reach a break-even threshold.

## [Results]{.underline}

The relationship between marketing expenditure and revenue is examined through both aggregate performance metrics (@tbl-roi-summary) and non-linear trend analysis (@fig-loess-trend).

As detailed in @tbl-roi-summary, the marketing campaign faces significant efficiency challenges. While the overall mean ROI stands at 12.7%, this figure is heavily skewed. The median ROI is negative (-6.9%), and 60.6% of all customer interactions have a negative ROI. This high prevalence of loss justifies the use of non-linear LOESS smoothing in @fig-loess-trend to identify the specific "efficiency windows" that maximize marginal returns.

```{r}
#| label: tbl-roi-summary
#| tbl-cap: "ROI Summary Statistics by Market Segment"
#| echo: false
#| warning: false
#| message: false


library(ggplot2)

library(kableExtra)
library(scales)

roi_table <- cleaned %>%
  filter(!is.na(roi)) %>%
  group_by(segment) %>%
  summarise(
    Customers          = n(),
    `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
    `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
    `Positive ROI (n)` = sum(roi > 0),
    `Negative ROI (n)` = sum(roi < 0),
    `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
    `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
    `Mean Purchase ($)` = paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2)),
    .groups = "drop"
  ) %>%
  arrange(segment)

# Add total row
total_row <- cleaned %>%
  filter(!is.na(roi)) %>%
  summarise(
    segment            = "ALL",
    Customers          = n(),
    `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
    `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
    `Positive ROI (n)` = sum(roi > 0),
    `Negative ROI (n)` = sum(roi < 0),
    `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
    `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
    `Mean Purchase ($)` = paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2))
  )

roi_summary <- bind_rows(roi_table, total_row)

roi_summary <- cleaned %>%
  filter(!is.na(roi)) %>%
  group_by(segment) %>%
  summarise(
    Customers          = n(),
    `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
    `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
    `Positive ROI (n)` = sum(roi > 0),
    `Negative ROI (n)` = sum(roi < 0),
    `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
    `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
    `Mean Purchase ($)`= paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2)),
    .groups = "drop"
  ) %>%
  bind_rows(
    cleaned %>%
      filter(!is.na(roi)) %>%
      summarise(
        segment            = "ALL",
        Customers          = n(),
        `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
        `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
        `Positive ROI (n)` = sum(roi > 0),
        `Negative ROI (n)` = sum(roi < 0),
        `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
        `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
        `Mean Purchase ($)`= paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2))
      )
  )

# 2. 判斷輸出格式並渲染
if (knitr::is_latex_output()) {
  # --- PDF 輸出設定 ---
  roi_summary %>%
    kable(
      format = "latex",
      booktabs = TRUE,
      align = c("l", rep("c", ncol(roi_summary) - 1))
    ) %>%
    kable_styling(
      latex_options = c("scale_down", "hold_position"),
      font_size = 8
    ) %>%
    column_spec(1, bold = TRUE) %>%
    row_spec(nrow(roi_summary), bold = TRUE, background = "#F2F2F2")
} else {
  # --- HTML 輸出設定 ---
  roi_summary %>%
    kable(
      format = "html",
      align = c("l", rep("c", ncol(roi_summary) - 1))
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    ) %>%
    column_spec(1, bold = TRUE) %>%
    row_spec(nrow(roi_summary), bold = TRUE) %>%
    scroll_box(width = "100%") # 預防 HTML 爆邊
}




```

```{r}
#| label: fig-loess-trend
#| fig-cap: "Non-linear smoothing (LOESS) identifying potential points of diminishing marginal returns in marketing spend."
#| echo: false
#| message: false
#| warning: false

ggplot(cleaned, aes(x = marketing_spend, y = purchase_amount, color = segment, fill = segment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", span = 0.9, se = TRUE, alpha = 0.1) + 
  labs(
    x = "Marketing Expenditure ($)",
    y = "Total Purchase Amount ($)",
    color = "Market Segment",
    fill = "Market Segment",
    caption = "Source: Campaign Performance Dataset 2024-2026"
  ) +
  theme_minimal()


```

The LOESS analysis in @fig-loess-trend reveals that revenue growth is not proportional to expenditure. Instead, distinct saturation thresholds appear across the three segments:

-   **Consumer Segment:** This segment is the most inelastic to marketing expenditure. The purple curve is flatter than the others, indicating a stable but lower marginal response to increased spending. While @tbl-roi-check shows a massive 85.0% ROI at low spend levels (\$0–\$100), efficiency decreases as spend increases. Because this segment lacks a strong growth trajectory, resources are likely better spent in the more responsive SMB and Enterprise segments.

-   **SMB Segment:** The SMB segment follows a classic pattern of diminishing marginal returns, where efficiency peaks early (steeper slope) with a 35.8% ROI in the \$0–\$100 bracket. However, as spending reaches the \$150–\$200 range, the curve flattens significantly. According to @tbl-roi-check , ROI for SMBs in the high-spend bracket (\>\$200) drops to -17.7%. This explains why 50% of this segment remains unprofitable.

-   **Enterprise Segment:** The Enterprise segment demonstrates a distinct minimum expenditure threshold, where growth only begins to accelerate once spending reaches approximately \$60. Between \$60 and \$150, Enterprise shows its highest marginal utility. However, a clear saturation point emerges at \$150, beyond which the curve flattens and returns turn negative (-23.2% for spend \>\$200).

```{r}
#| label: tbl-roi-check
#| echo: false
#| message: false

# 1. Calculate granular ROI for ALL segments
roi_check_all <- cleaned %>%
  filter(!is.na(roi)) %>%
  mutate(spend_bracket = cut(marketing_spend, 
                             breaks = c(0, 100, 200, Inf),
                             labels = c("Low ($0-100)", "Mid ($100-200)", "High ($200+)"))) %>%
  group_by(segment, spend_bracket) %>%
  summarise(
    avg_roi = mean(roi, na.rm = TRUE),
    sample_size = n(),
    .groups = "drop"
  ) %>%
  # Convert to percentage for display
  mutate(avg_roi_pct = scales::percent(avg_roi, accuracy = 0.1))

# 2. Display the table
roi_check_all %>%
  select(segment, spend_bracket, avg_roi_pct, sample_size) %>%
  knitr::kable(
    caption = "Detailed ROI Analysis by Segment and Spend Bracket",
    col.names = c("Segment", "Spend Bracket", "Avg ROI (%)", "Sample Size (n)"),
    align = "llcc"
  ) %>%
  kableExtra::kable_styling(bootstrap_options = c("condensed", "striped", "hover"), 
                            full_width = FALSE) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top") # Merges segment names for clarity


```

The granular data in @tbl-roi-check confirms that for all three segments, spending over \$200 leads to guaranteed negative returns (ranging from -17.7% to -23.2%). To maximize total revenue and move the campaign toward profitability, the following budget adjustments are recommended:

-   **Exploit the Efficiency Zone (\$0–\$100):** For the Consumer and SMB segments, marketing expenditure below \$100 yields the highest returns, with average ROIs of 85.0% and 35.8%, respectively. Management should focus on increasing the volume of customers within this low-spend bracket, as it represents the most cost-effective growth opportunity.

-   **Target the Enterprise "Growth Window" (\$60–\$150):** The Enterprise segment requires a minimum "entry spend" to trigger returns; investments below \$60 result in heavy losses (-47.8% ROI). Spending should be strictly targeted within the \$60–\$150 window to capture the positive ROI (10.3%) seen in the mid-spend bracket before saturation begins.

-   **Enforce Universal Spending Caps (\$200):** Across all segments, a hard cap of \$200 per customer should be implemented. Expenditure beyond this point consistently erodes total campaign profit, indicating that these customers are over-saturated and no longer responsive to additional marketing pressure.
