---
title: "ETC5513_Assignment1"
author: "Sinping CHAN"
editor: visual
format:
  html:
    theme: cosmo       
    toc: true
    code-fold: true      
    embed-resources: true 
  pdf:
    documentclass: article
    toc: true
    number-sections: true
    colorlinks: true
execute:
  echo: false         
  warning: false
  message: false
---

```{r}
#| label: quicklook
#| include: false
#| warning: false
#| message: false

library(tidyverse)
raw <- read_csv("data/35851937_Sin Chan.csv")

head(raw)
names(raw)


```

## Introduction

> In an increasingly competitive e-commerce landscape, quantifying the return on personalized marketing investments is essential for driving sustainable revenue growth. This research examines whether higher expenditure on advertisements, emails, and coupons during the previous quarter’s personalized campaign led to increased individual purchase amounts.\
> \
> By analyzing these relationships at the customer level, the study evaluates the effectiveness of variable marketing spend across different channels.\

## Research Question

> The main goal of this report is to determine if increasing investment in a specific channel results in a linear increase in revenue or if data shows distinct points of diminishing return for each category. The study analyzes these relationships in order to provide actionable insights that can be used to optimize future marketing budgets for diverse communication channels.

## Dataset Introduction

```{r}
#| label: data-cleansing
#| include: false
#| warning: false
#| message: false

# ============================================================
#  Data quality issues addressed:
#    1. Mixed date formats  → standardised to ISO 8601 (YYYY-MM-DD)
#    2. Missing signup_date  (9 rows)  → kept as NA
#    3. Missing purchase_amount (18 rows) → kept as NA; flag added
#    4. notes column: NaN → NA; meaningful values retained
#    5. segment → ordered factor (Consumer < SMB < Enterprise)
#    6. Derived features: days_since_signup, roi, contacted_flag,
#                         returned_flag, has_purchase
# ============================================================

# ── 0. PACKAGES ─────────────────────────────────────────────
library(dplyr)
library(lubridate)
library(stringr)
library(readr)

# ── 1. LOAD RAW DATA ────────────────────────────────────────

raw <- read_csv(
  "data/35851937_Sin Chan.csv",   # adjust path if needed
  col_types = cols(
    customer_id      = col_character(),
    signup_date      = col_character(),   # read as text; parsed below
    marketing_spend  = col_double(),
    visits           = col_integer(),
    purchase_amount  = col_double(),
    segment          = col_character(),
    notes            = col_character()
  ),
  na = c("", "NA", "N/A", "na", "n/a")   # normalise common NA strings on read
)

cat("Raw data loaded:", nrow(raw), "rows x", ncol(raw), "cols\n")


# ── 2. DATE PARSING HELPER ──────────────────────────────────
# The column contains (at least) four formats:
#   YYYY-MM-DD     e.g. "2025-06-24"
#   DD/MM/YYYY     e.g. "22/05/2025", "27/10/2024"
#   MM/DD/YYYY     e.g. "05/10/2024"  ← ambiguous; resolved via context*
#   Mon DD, YYYY   e.g. "Feb 05, 2025", "Oct 10, 2024"
#
# * DD/MM vs MM/DD is ambiguous when day <= 12. Inspection of the data
#   confirms the supplier locale uses DD/MM/YYYY throughout; that format
#   is tried before MM/DD/YYYY so it takes precedence.

parse_mixed_date <- function(x) {
  formats <- c(
    "%Y-%m-%d",    # ISO  – try first (most common)
    "%d/%m/%Y",    # DD/MM/YYYY
    "%m/%d/%Y",    # MM/DD/YYYY (fallback)
    "%b %d, %Y",   # Mon DD, YYYY  (abbreviated month)
    "%B %d, %Y"    # Month DD, YYYY (full month name)
  )

  result <- as.Date(rep(NA_character_, length(x)))

  for (fmt in formats) {
    needs_filling <- is.na(result) & !is.na(x)
    if (!any(needs_filling)) break
    parsed <- suppressWarnings(as.Date(x, format = fmt))
    result[needs_filling & !is.na(parsed)] <- parsed[needs_filling & !is.na(parsed)]
  }
  result
}


# ── 3. CLEAN ────────────────────────────────────────────────

reference_date <- as.Date("2026-03-31")   # today; used for derived fields

cleaned <- raw %>%

  # 3a. Parse signup_date
  mutate(
    signup_date = parse_mixed_date(signup_date)
  ) %>%

  # 3b. notes: already NA-normalised on read; standardise case
  mutate(
    notes = str_to_lower(str_trim(notes))   # "Contacted" → "contacted"
  ) %>%

  # 3c. segment → ordered factor
  mutate(
    segment = factor(
      segment,
      levels  = c("Consumer", "SMB", "Enterprise"),
      ordered = TRUE
    )
  ) %>%

  # 3d. Derived features
  mutate(
    # Days between signup and reference date (NA if signup unknown)
    days_since_signup = as.integer(reference_date - signup_date),

    # Return on marketing investment; NA if either value is missing
    roi = ifelse(
      !is.na(purchase_amount) & marketing_spend > 0,
      round((purchase_amount - marketing_spend) / marketing_spend, 4),
      NA_real_
    ),

    # Boolean flags derived from notes
    contacted_flag  = (!is.na(notes) & notes == "contacted"),
    returned_flag   = (!is.na(notes) & notes == "returned"),
    no_response_flag = (!is.na(notes) & notes == "no response"),

    # Flag rows with missing purchase_amount for downstream imputation decisions
    purchase_missing = is.na(purchase_amount)
  )


# ── 4. VALIDATION CHECKS ────────────────────────────────────

cat("\n── Cleaned data summary ──────────────────────────────\n")
cat("Rows          :", nrow(cleaned), "\n")
cat("Columns       :", ncol(cleaned), "\n\n")

cat("Missing values per column:\n")
print(colSums(is.na(cleaned)))

cat("\nSegment distribution:\n")
print(table(cleaned$segment))

cat("\nnotes distribution:\n")
print(table(cleaned$notes, useNA = "ifany"))

cat("\nsignup_date range (non-missing):\n")
cat("  Min:", format(min(cleaned$signup_date, na.rm = TRUE)), "\n")
cat("  Max:", format(max(cleaned$signup_date, na.rm = TRUE)), "\n")

cat("\npurchase_amount summary:\n")
print(summary(cleaned$purchase_amount))

cat("\nmarketing_spend summary:\n")
print(summary(cleaned$marketing_spend))

cat("\nROI summary:\n")
print(summary(cleaned$roi))


# ── 5. EXPORT ───────────────────────────────────────────────

write_csv(cleaned, "35851937_Sin_Chan_cleaned.csv", na = "")

```

### Data Cleaning

> To ensure the dataset was ready for reproducible analysis, a cleaning procedure was performed. First, the `signup_date` column was standardized into a consistent ISO 8601 format to resolve various date string discrepancies. Second, inconsistent missing value indicators (such as 'N/A' or 'na') were unified into standard R `NA` values across all columns, particularly for `purchase_amount` and `notes`. Textual entries in the `notes` column were trimmed of extra whitespace and converted to lowercase to maintain string consistency. Finally, the `segment` variable was reclassified as an ordered factor to correctly represent the hierarchy of business tiers, ensuring that the statistical summaries reflect the intended business logic.

### Data Description

> This dataset records the engagement of `r nrow(cleaned)` customers across 3 market segments (Consumer, SMB, and Enterprise) from September 2024 to February 2026. The data measures key performance metrics including individual marketing expenditures, total purchase amounts, and platform visit frequency. The processed dataset consists of `r ncol(cleaned)` variables, ranging from numerical financial figures (e.g., `marketing_spend`, `roi`) to categorical classifications and logical flags used for tracking customer behavior.
>
> The dataset includes the following 6 variable types:
>
> -   **Numeric**: `marketing_spend`, `purchase_amount`, and `roi`.
>
> -   **Integer**: `visits` and `days_since_signup`.
>
> -   **Ordered Factor**: `segment` (Consumer \< SMB \< Enterprise).
>
> -   **Date**: `signup_date`.
>
> -   **Character**: `customer_id` and `notes`.
>
> -   **Logical**: `contacted_flag`, `returned_flag`, `no_response_flag`, and `purchase_missing`

```{r}
#| label: fig-spend-vs-purchase
#| echo: false
#| warning: false

library(ggplot2)

ggplot(cleaned, aes(x = marketing_spend, y = purchase_amount, color = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Figure 1: Marketing Spend vs Purchase Amount by Segment",
    subtitle = "Analysis of correlation across different business tiers",
    x = "Marketing Spend ($)",
    y = "Purchase Amount ($)",
    color = "Segment"
  ) +
  theme_minimal()

```

```{r}
#| label: tbl-segment-summary
#| echo: false
#| message: false

library(dplyr)
library(knitr)

# 建立按 segment 分組的統計表
segment_summary <- cleaned |>
  group_by(segment) |>
  summarise(
    `Avg Spend` = mean(marketing_spend, na.rm = TRUE),
    `Avg Purchase` = mean(purchase_amount, na.rm = TRUE),
    `Total Visits` = sum(visits, na.rm = TRUE),
    `Missing Rate (%)` = mean(purchase_missing) * 100
  ) |>
  mutate(across(where(is.numeric), ~round(., 2)))

# 在 kable 函數內直接加入 caption
kable(segment_summary, 
      caption = "Table 1: Summary Statistics of Campaign Performance by Segment")
```

As detailed in @tbl-segment-summary, the average marketing spend and purchase behavior vary across 3 segments, with the Enterprise segment showing the highest average marketing spend at \$174.73, while the Consumer segment contributed the vast majority of platform traffic with 511 total visits.\
Furthermore, @fig-spend-vs-purchase illustrates the linear relationship between spending and revenue.

------------------------------------------------------------------------

## Results

```{r}
# ── PACKAGES ────────────────────────────────────────────────
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ══════════════════════════════════════════════════════════════
# ROI SUMMARY TABLE
# ══════════════════════════════════════════════════════════════

#| label: tbl-roi-summary
#| echo: false
#| warning: false
#| message: false

roi_table <- cleaned %>%
  filter(!is.na(roi)) %>%
  group_by(segment) %>%
  summarise(
    Customers          = n(),
    `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
    `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
    `Positive ROI (n)` = sum(roi > 0),
    `Negative ROI (n)` = sum(roi < 0),
    `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
    `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
    `Mean Purchase ($)` = paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2)),
    .groups = "drop"
  ) %>%
  arrange(segment)

# Add total row
total_row <- cleaned %>%
  filter(!is.na(roi)) %>%
  summarise(
    segment            = "ALL",
    Customers          = n(),
    `Mean ROI (%)`     = paste0(round(mean(roi) * 100, 1), "%"),
    `Median ROI (%)`   = paste0(round(median(roi) * 100, 1), "%"),
    `Positive ROI (n)` = sum(roi > 0),
    `Negative ROI (n)` = sum(roi < 0),
    `% Negative ROI`   = paste0(round(mean(roi < 0) * 100, 1), "%"),
    `Mean Spend ($)`   = paste0("$", round(mean(marketing_spend), 2)),
    `Mean Purchase ($)` = paste0("$", round(mean(purchase_amount, na.rm = TRUE), 2))
  )

roi_summary <- bind_rows(roi_table, total_row)

roi_summary %>%
  kable(
    caption  = "ROI Summary Statistics by Segment",
    align    = c("l", rep("c", ncol(roi_summary) - 1)),
    booktabs = TRUE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE,
    position          = "left"
  ) %>%
  column_spec(1, bold = TRUE) %>%
  column_spec(6, bold = TRUE,
              color = ifelse(
                as.numeric(gsub("%", "", roi_summary$`% Negative ROI`)) > 60,
                "red", "black"
              )) %>%
  row_spec(nrow(roi_summary), bold = TRUE, background = "#f5f5f5")

```

```{r}
#| label: fig-diminishing-returns
#| fig-cap: "Non-linear smoothing (LOESS) identifying potential points of diminishing marginal returns in marketing spend."
#| echo: false
#| message: false
#| warning: false

library(ggplot2)

ggplot(cleaned, aes(x = marketing_spend, y = purchase_amount, color = segment, fill = segment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", span = 0.9, se = TRUE, alpha = 0.1) + # 呢度一定要有 + 號
  labs(
    title = "Marketing Spend vs. Purchase Revenue",
    subtitle = "Analysis of Diminishing Returns across Customer Segments",
    x = "Marketing Expenditure ($)",
    y = "Total Purchase Amount ($)",
    color = "Market Segment",
    fill = "Market Segment",
    caption = "Source: Campaign Performance Dataset 2024-2026"
  ) +
  theme_minimal()
```

The LOESS analysis in **\@fig-diminishing-returns** reveals distinct saturation thresholds across segments, addressing the research question on budget optimization:

-   **SMB Segment**: This group demonstrates a strong initial response, exhibiting peak efficiency with a steep slope between **\$0–\$100**, suggesting a high Return on Investment (ROI) for early-stage spending. A temporary saturation point emerges around **\$150–\$200**, where the curve noticeably flattens, marking the threshold where diminishing returns settle in. Beyond \$200, the segment demonstrates a secondary growth phase, suggesting a **two-stage response curve** rather than a simple linear relationship. SMB investment should therefore be prioritised within the **\$0–\$150 high-efficiency window**.

<!-- -->

-   **Enterprise Segment**: This group demonstrates the most efficient initial response. The data shows that Enterprise clients in this dataset have a minimum observed marketing spend of approximately **\$60**, beyond which a steep and efficient growth trajectory is evident. However, a clear saturation point emerges at approximately **\$150**, beyond which the curve flattens and no longer outperforms other segments. Budget allocation for Enterprise clients should therefore be concentrated within the **\$60–\$150 active growth window** to maximise marginal returns.

<!-- -->

-   **Consumer Segment**: This segment is **relatively inelastic** to marketing expenditure and appears to be the least responsive to specific spending fluctuations. The curve remains relatively linear but flatter throughout the range, indicating a **stable but lower marginal utility** compared to the B2B segments. Due to this lack of significant marginal utility windows, resources could be more effectively **reallocated** to the optimised **SMB** and **Enterprise** windows to maximise total revenue.

To maximise total revenue, priority should be given to **SMB investments up to \$150** and **Enterprise spending concentrated within the \$60–\$150 window** before diminishing returns occur
